
# rlist 


[rlist](https://renkun-ken.github.io/rlist-tutorial/Getting-started/Quick-overview.html)
ß
```{r}
# install.packages('rlist')
library(rlist)
library(tidyverse)
# install.packages('pipeR')
library(pipeR)

url <- "https://renkun-ken.github.io/rlist-tutorial/data/people.json"

people <- list.load(url)

people %>>%
  list.filter(Expertise$R >= 1 & Expertise$Python >= 1) %>>%
  list.class(Interests) %>>%
  list.sort(-length(.)) %>>%
  list.take(3) %>>%
  list.map(. %>>% list.table(Age))
```


- We filter people list by the prerequisites: Using both R and Python for at least one year.

-  Then we class all people by their interests, that is, we create a big list of possible interest values, each is a nested list of all people whose interests contain that value.

- Then we sort in descending these interest classes by the number of people who has the corresponding interest.

- Then we pick out the top 3 interests with most people.
- Then we map each interest class to a table of ages in that class.
-  The output is exactly the answer to the question.


| Name   |	Age	| Interests	              | Expertise             |
| -------|------|-------------------------|-----------------------|
| Ken    |	24	| reading, music, movies	| R:2, C#:4, Python:3   |
| James  |	25	| sports, music	          | R:3, Java:2, C++:5    |
| Penny	 |  24	| movies, reading	        | R:1, C++:4, Python:2  |



using a csv would not work for all of the data
using a JSON file would work but is messy


rlist package imports jsonlite package to read/write JSON data.

Another file format that is also widely used is `YAML`. The following text is a YAML format representation (stored here) of the non-tabular data:
rlist also imports yaml package to read/write YAML data.

```{}
# YAML file 

- Name: Ken
  Age: 24
  Interests:
  - reading
  - music
  - movies
  Expertise:
    R: 2
    CSharp: 4
    Python: 3
    
- Name: James
  Age: 25
  Interests:
  - sports
  - music
  Expertise:
    R: 3
    Java: 2
    Cpp: 5
    
- Name: Penny
  Age: 24
  Interests:
  - movies
  - reading
  Expertise:
    R: 1
    Cpp: 4
    Python: 2
```



# mapping

use the small JSON data

```{r}
j = "https://renkun-ken.github.io/rlist-tutorial/data/sample.json"

small_json = list.load(file= j, type = 'json')

str(small_json)
```



To extract the name of each people (list element), traditionally we can call `lapply()`

```{r}
# people, prints out 500 names from the Name list

lapply(small_json, function(x) {
  x$Name
})
```



Using rlist's list.map() to extract all names in Name

```{r}
list.map(small_json, Name)
```

The following code maps each list member to the sum of years of the programming languages they use.

```{r}
list.map(small_json, Expertise)
```

sum up each person's experience by all coding expertise 

```{r}
list.map(small_json, sum( as.numeric(Expertise) )  )
```

map person name with experience range, James has most coding experience

```{r}
list.map(small_json, list(Name=Name, range=range(as.numeric(Expertise))))
```


### list.map shorthand

In some cases we need to refer to the item itself, or its index in the list, or even its name. In the expression, `.` represents the item itself, `.i` represents its index, and `.name` represents its name.

- a 3 becomes 4
- b 2 becomes 3
- c 1 becomes 2

```{r}
nums =  c(a=3, b=2, c=1)

list.map(nums, . + 1)

list.map(nums, paste0("name: ", .name))
```


### mapv

get a vector instead of a list returned

```{r}
list.mapv(small_json, Age)

list.mapv(small_json, Interests)

list.mapv(small_json, Expertise)
```


### list.select

```{r}
list.select(small_json, Name, Expertise)
```

```{r}
list.select(small_json, Name, Age, CodingExp= length(Expertise))
```



### list.iter

just get a value

```{r}
list.iter(small_json, cat(Name, ":", Age, "\n"))
```

### list.maps

can use more than one list

```{r}
L1 = list(p1= list(x=1, y=2), 
          p2= list(x=3, y=4), 
          p3= list(x=1, y=3) 
          )

L2 = list(2, 3, 5)

list.maps(a$x * b + a$y, a= L1, b=L2)
```



## Filter list

list.filter() filters a list by an expression that returns TRUE or FALSE.

```{r}
list.filter(small_json, Age >= 25)
```

```{r}
library(pipeR)

small_json %>>%
  list.filter(Age >= 25) %>>%
  list.mapv(Name)
```


## List find

```{r}
people %>>%
  list.find(Age >= 25,n = 10) %>>%
  list.mapv(Name)


# find index number location
list.findi(people, Age >= 23, 2)
```

List take

```{r}
list.take(5:10, 3)
```

list skip

```{r}
list.skip(1:10, 3)
```

list.count
```{r}
list.count(people, "R" %in% names(Expertise))
```


list.match

```{r}
data = list(a1= 3, a3= 7, c= 9)

list.match(data, "a[13]")
```


list subset

```{r}
small_json %>>% 
  subset(Age >= 24, Name)
```


```{r}
small_json %>>%
  list.select(Name, Age) %>>%
  list.stack
```


Suppose we find that the age of each people is mistakenly recorded, say, 1 year less than their actual ages, respectively, we need to update the original data by refresh the age of each element.

```{r}
small_json %>>%
  list.update(Age = Age + 1) %>>%
  list.select(Name, Age) %>>%
  list.stack
```



list.order

```{r}
list.order(small_json, Age)

list.order(small_json, (Expertise$R))
```


list sort
```{r}
small_json %>>%
  list.sort(Age) %>>%
  list.select(Name, Age)
```

```{r}
small_json %>>%
  list.sort((Expertise$R)) %>>%
  list.select(Name, R = Expertise$R)
```



list cases
```{r}
list.cases(small_json, Interests)

list.cases(people, names(Expertise))
```


```{r}
small_json %>>%
  list.class(Interests) %>>%
  list.map(. %>>% list.mapv(Name))
```


```{r}
list.common(small_json, names(Expertise))
```


list.table

table() builds a contingency table of the counts at each combination of factor levels using cross-classifying factors. list.table() is a wrapper that creates a table in which each dimension results from the values for an expression.

The function is very handy to serve as a counter. The following examples shows an easy way to know the remainders and the number of integers from 1 to 1000 when each is divided by 3.

```{r}
list.table(1:1000, . %% 3)
```



list.join

```{r}
newinfo = list(
    list(Name="Ken", Email="ken@xyz.com"),
    list(Name="Penny", Email="penny@xyz.com"),
    list(Name="James", Email="james@xyz.com")
    )


list.join(small_json, newinfo, Name)
```



Search all elements equal to Ken recursively.

```{r}
friends <- list.load("https://renkun-ken.github.io/rlist-tutorial/data/friends.json")


list.search(friends, . == "Ken")
```


## list.parse

convert an object to a list, convert dataframe or a matrix to a list

```{r}
df = data.frame(
  names = c("Linux","Mac","Windows"),
  age = c(58,57,48),
  stringsAsFactors = F
)

list.parse( df)
```

JSON data

```{r}
js = '[
    {"name": "Apple", "age": 67},
    {"name": "Windows", "age": 78},
    {"name": "Linux", "age": 44}
  ]'

str( list.parse(js, "json") ) 
```

YAML data

```{r}
yam = "
p1:
  name: Apple
  age: 167
p2:
  name: Windows
  age: 89
p3:
  name: Linux
  age: 56
"

parsed_yaml = list.parse(yam, "yaml") 
str( parsed_yaml ) 
```



## list.stack

the opposite of list.parse() on a dataframe, creates a dataframe
Take the parsed yaml data and put in list.stack

```{r}
list.stack( parsed_yaml)
```

## list.load & list.save

this looks at file extension and determines which loader to use

```{r}
j = "https://renkun-ken.github.io/rlist-tutorial/data/sample.json"

small_json = list.load(file= j, type = 'json')
```



## list.append

```{r}
library(rlist)

list.append(list(a=100, b= 200, c= 300))

```

list.prepend

```{r}
list.prepend( list(alpha = 10, beta= 20, ceta= 30))
```
```{r}
list.append( c(a=1, b= 2), c=3)
```





## list.zip

```{r}
str( list.zip(a= c(1,2,3), b= c(10,20,30)) )
```


```{r}
str( list.zip(x= list(1,"x"), y= list("y", 2)) )
```



## list.rbind

```{r}
scores = list(
  score1 = c(10,20,30),
  score2 = c(100, 200, 300),
  score3 = c(1e3, 2e3, 3e3)
)

list.rbind( scores)
```

```{r}
list.cbind( scores )
```


a list of lists => matrix of lists

```{r}
scores3 = list(
  score1.1 = list(150, 250, 350),
  score2.1 = list(1150, 2150, 3150),
  period = list('one','two','three')
)

scores3_r = list.rbind( scores3)
scores3_r
```

```{r}
scores3_r[1,1]
```



## list.stack

to create a dataframe from lists, helpful for non tabular data

```{r}
nontab = list(
  list(
    type="A",
    value= 110
  ),
  list(
    type="B",
    value= 90
  )
)

list.stack( nontab)
```


grab data then stack columns we want

```{r}
list.load("https://renkun-ken.github.io/rlist-tutorial/data/sample.json") %>>%
  list.select(Name, Age) %>>%
  list.stack
```


## list.flatten

```{r}
d = list(
  list(
    a= 10,
    b= 20
  ),
  list(
    c(c= 245,
      d= list(w= 899,
              h= 1788 ))
  )
)

# d
str(d)
```

```{r}
list.flatten(d)
```



## lambda expressions

```{r}
x = list(
  a= c(w= 10, d= 20 ),
  b = c(w= 100, d= 200)
)

list.map(x, sum(d))
```

```{r}
list.map(x,  w*d)
```

```{r}
list.map(x, w + .i)
```


```{r}
list.map(x, .name)
```




## list enviro

```{r}
people <- list.load("https://renkun-ken.github.io/rlist-tutorial/data/sample.json")

ppl = List(people)

ppl$map(Name)
```





# GitHub data

Website data is in JSON format

```{r}
repos = "https://api.github.com/users/hadley/repos?per_page=100&page=%d" 

repos = repos %>>%
  sprintf(1:2) %>>%
  list.load("json") %>>%
  list.ungroup
```



```{r}
list.count( repos)
```

```{r}
repos %>% 
  list.table( fork )
```

find out the language types of the repos

```{r}
repos %>% 
  list.filter( !is.null( language)) %>% 
  list.table( language ) %>% 
  list.sort( - .) # descending 
```

table of language by fork 

```{r}
repos %>% 
  list.table( language, fork)
```




```{r}
library(ggplot2)

top_10= repos %>>%
  list.filter(!fork, language == "R") %>>%
  list.names(name) %>>%
  list.mapv(stargazers_count) %>>%
  list.sort(-.) %>>%
  list.take(10)  %>>%
  print %>>%
  barplot(main = "Hadley's top 10 R repos with most stargazers")

str(top_10)
```



```{r}
repos %>>%
  list.filter(has_issues, !fork, language == "R") %>>%
  list.names(name) %>>%
  list.mapv(open_issues) %>>%
  list.sort(-.) %>>%
  list.take(10) %>>%
  print %>>%
  barplot(main = "Hadley's top 10 R repos with most open issues")
```



```{r}
repos %>>%
  list.filter(any(grepl("plot|vis", c(name, description)))) %>>%
  list.sort(-stargazers_count) %>>%
  list.mapv(name)
```

```{r}
repos %>>%
  list.select(stargazers_count, watchers_count, forks_count) %>>%
  list.stack %>>%
  colSums
```




